\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-08 cfhtlens.tex
Added Fabian's text Accepted SJ's changes (thx SJ)

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to their Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}. This can be compared to the full five-parameter fit from \citet{kilbinger/etal:2013} who find $S_8=\sigma_8(\Omega_m/0.3)^{0.6} = 0.74 \pm 0.03$.   We argue that this reduction in the relevant error by a factor of $\sim 2$, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} and are unable to confirm the results quoted by \citet{kitching/etal:2016}. 
%{\bf TBD:  The results that follow use the exact Planck 2015 cosmology.  We just noticed footnote 5 in Kitching et al that shows that they've actually used an approximate Planck 2015 cosmology - currently re-running with this alternative which is unlikely to change these results significantly}.  
We find the lowest conditional value of $\sigma_8 = 0.7461 \pm 0.0231$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.7487 \pm 0.0232$ for the `Baseline Limber Flat' (LF) case. Comparing the fiducially used `Extended Standard Flat' (ESt) case with the preferred `Extended Limber Spherical' (ELS) case, the constraints are identical at the $0.01\sigma$ level.
Therefore, as expected from the comparison in Figure~\ref{fig:Cl_xi}, and given the low signal-to-noise CFHTLenS measurements on large scales, we find that the different combinations of assuming flat or spherical sky, baseline or extended Limber, make little difference to the cosmological parameter constraints, even in this `one-parameter' analysis.

Our Limber-case findings are in disagreement with \citet{kitching/etal:2016} who find a shift of $\Delta \sigma_8 = 0.007$ between their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) and their `Limber' analysis (which we believe corresponds to the Extended Limber spherical (ELS) case in Table~\ref{tab:Tl_nu}).  
\shahab{(I believe Tom in the latest email claimed the 0.007 number is between ESt and ELS)}

We note that it appears \citet{kitching/etal:2016} have chosen to use their `Flat' approximation in their `Full' non-Limber analysis rather than a full spherical-sky solution.  The rationale is based on the Hankel transform in equation~\ref{eqn:xiGG} which is derived assuming a flat-sky approximation.  As shown in Figure~\ref{fig:Cl_xi}, however their chosen version of a flat-sky approximation serves to exacerbate the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution. We do not find this to be the case when considering the standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.

\subsection{Comparison of two-point shear statistics; the two-point correlation function, mass aperture statistic and COSEBIs}
\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range different two-point shear statistics including the shear correlation function, $\xi_\pm$, the aperture-mass dispersion, $M_ {\rm ap}^2$ \citep{schneider/etal:1998}, and the COSEBIs \citep{schneider/etal:2010}.  Figure~\ref{fig:filters} shows the integrands of these statistics 
% (i.e. the weights placed on each $\ell$ scale),  
for two cases, normalised to their maximum value.  The integrands are of the form $\ell F(\ell) P(\ell)$, where $F(\ell)=J_0(\ell)$ for $\xi_+$, $F(\ell)= U^2_\theta(\ell)$ for  $M_ {\rm ap}^2$ \citep[see][]{schneider/etal:1998} and $F(\ell)=W_n(\ell)$ for COSEBIs \citep[see][]{schneider/etal:2010}. For $\xi_\pm$, $P(\ell)$ is equal to the sum of the E and B-mode convergence power spectra. This is one of the main motivators for the development of aperture-mass dispersion. However, $M_ {\rm ap}^2$ statistics is a lossy conversion and is biased by small angular separations,
where blending of galaxies makes shear measurement challenging/impossible \citep{Kilbinger/etal:2006}. As a result, COSEBIs was developed which tackles both shortcomings of $M_ {\rm ap}^2$ statistics.
The right panel shows the integrand\footnote{Note that the lower left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum angular separation of $100'$. However, in \citet{kilbinger/etal:2013}, the data used in \citet{kitching/etal:2016}, $\theta$ is between $0.8'$ and $350'$.} of $\xi_+$ for $\theta=100'$ and $\theta=350'$. The middle panel in Figure~\ref{fig:filters} shows COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode, as the higher modes generally probe larger $\ell$-modes.  Finally, the left panel shows the integrands of aperture mass dispersion statistics, for the same two maximum angular ranges.

@@

>>>>

\citet{kitching/etal:2016} choose to only vary $\sigma_8$ in their analysis, fixing all other parameters to Planck values. They find a conditional value of $\sigma_8 = 0.789 \pm 0.015$ for their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) of the data from \citet{kilbinger/etal:2013}. This can be compared to the full five-parameter fit from \citet{kilbinger/etal:2013} who find $S_8=\sigma_8(\Omega_m/0.3)^{0.6} = 0.74 \pm 0.03$.   We argue that this reduction in the relevant error by a factor of $\sim 2$, could easily mislead, making it appear that the impact of removing all approximations from the theoretical analysis is more significant for current surveys than it is in reality.  Nevertheless, we repeated the `one-parameter' analysis for the full range of Limber cases listed in Table~\ref{tab:Tl_nu} and are unable to confirm the results quoted by \citet{kitching/etal:2016}.   We find the lowest conditional value of $\sigma_8 = 0.746 \pm 0.023$ for the `Standard' (KSt) case and the highest conditional value of $\sigma_8 = 0.749 \pm 0.023$ for the `Baseline Limber Flat' (LF) case, a maximum change of $\Delta \sigma_8 = 0.0026 = 0.11 \sigma$.  These findings are in disagreement with \citet{kitching/etal:2016} who find a shift of $\Delta \sigma_8 = 0.007 = 0.5\sigma$ between their `standard' analysis (row 1 in Table~\ref{tab:Tl_nu}) and their `Limber' analysis (which we believe corresponds to the Extended Limber spherical (ELS) case in Table~\ref{tab:Tl_nu}).

The most relevant case to consider is differences between the fiducially used `Extended Standard Flat' (ESt) case with the preferred `Extended Limber Spherical' (ELS) case.  Here we find that the constraints are identical at the $0.01\sigma$ level.  As expected from the comparison in Figure~\ref{fig:Cl_xi}, and given the low signal-to-noise CFHTLenS measurements on large scales, we conclude that the different combinations of assuming flat or spherical sky, baseline or extended Limber, make little difference to the cosmological parameter constraints, even in a `one-parameter' analysis.

We note that \citet{kitching/etal:2016} have chosen to use their `Flat' approximation in their `Full' non-Limber analysis rather than a full spherical-sky solution.  The rationale is based on the Hankel transform in equation~\ref{eqn:xiGG} which is derived assuming a flat-sky approximation.  As shown in Figure~\ref{fig:Cl_xi}, however their chosen version of a flat-sky approximation serves to exacerbate the tension in theoretical models when a flat-sky solution is compared with the spherical-sky solution. We do not find this to be the case when considering the standard $T_\ell = 1$ flat-sky approximation that has been used in the majority of cosmic shear studies to date.

\subsection{Comparison of two-point shear statistics; the two-point correlation function, mass aperture statistic and COSEBIs}
\citet{kilbinger/etal:2013} present a detailed comparison of cosmological constraints obtained from a range different two-point shear statistics including the shear correlation function, $\xi_\pm$, the aperture-mass dispersion, $M_ {\rm ap}^2$ \citep{schneider/etal:1998}, and the COSEBIs \citep{schneider/etal:2010}.  Figure~\ref{fig:filters} shows the integrands of these statistics 
% (i.e. the weights placed on each $\ell$ scale),  
for two cases, normalised to their maximum value.  The integrands are of the form $\ell F(\ell) P(\ell)$, where $F(\ell)=J_0(\ell)$ for $\xi_+$, $F(\ell)= U^2_\theta(\ell)$ for  $M_ {\rm ap}^2$ \citep[see][]{schneider/etal:1998} and $F(\ell)=W_n(\ell)$ for COSEBIs \citep[see][]{schneider/etal:2010}. For $\xi_\pm$, $P(\ell)$ is equal to the sum of the E and B-mode convergence power spectra. This is one of the main motivators for the development of aperture-mass dispersion. However, $M_ {\rm ap}^2$ statistics is a lossy conversion and is biased by small angular separations,
where blending of galaxies makes shear measurement challenging/impossible \citep{kilbinger/etal:2006}. As a result, COSEBIs was developed which tackles both shortcomings of $M_ {\rm ap}^2$ statistics.
The right panel shows the integrand\footnote{Note that the lower left panel of Fig. 1 in \citet{kitching/etal:2016} shows an integrated form of this function for a maximum angular separation of $100'$. However, in \citet{kilbinger/etal:2013}, the data used in \citet{kitching/etal:2016}, $\theta$ is between $0.8'$ and $350'$.} of $\xi_+$ for $\theta=100'$ and $\theta=350'$. The middle panel in Figure~\ref{fig:filters} shows COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode, as the higher modes generally probe larger $\ell$-modes.  Finally, the left panel shows the integrands of aperture mass dispersion statistics, for the same two maximum angular ranges.

@@

\subsection{On the quadratic estimator}

\citet{kitching/etal:2016} state that `the derivative of a Limber-approximated power spectrum is not equal to the Limber-approximation of the derivative'. Based on this statement they conclude that `statistics that use Limber-approximated quadratic estimators to measure the cosmic shear power spectrum are not accurate' and refer explicitly to the quadratic estimator implementation used in \citet{Koehlinger2016} to measure the cosmic shear power spectrum from CFHTLenS. In this quadratic estimator implementation, however, the Limber approximation does not enter. The algorithm was originally presented by \citet{Hu2001} in the context of cosmic shear and has also been applied to other shear data sets (e.g. \citealt{Brown2003, Heymans2005, Lin2012}). Below we briefly summarize this quadratic estimator algorithm and emphasize that the Limber approximation does not enter in there.

The aim of the method is to find the best-fitting power spectra assuming a Gaussian likelihood function $\mathcal{L}$. The likelihood depends on the shear data vector and its covariance $\mathbf{C}$ which consists of the sum of the shear correlation matrix $\mathbf{C}^\mathrm{sig}$ and the noise matrix $\mathbf{C}^\mathrm{noise}$. The shear correlation matrix can be written as a linear combination of piece-wise constant band powers $\mathcal{B}$ approximating the Fourier transforms of the shear correlations, i.e. their power spectra (cf. equations~3,~10 in \citealt{Koehlinger2016}). The best-fitting power spectra are found by iteratively stepping through a Newton--Raphson method in order to find the root of $\diff \mathcal{L}/ \diff \mathcal{B} = 0$ \citep{Bond1998, Seljak1998}. In the subsequent equations of the Newton--Raphson method only derivatives of the full covariance matrix enter, where the derivatives are taken with respect to band powers\footnote{i.e. $\partial \mathbf{C} / \partial \mathcal{B}_A$, where $A$ is a superindex over the type of the band, band-power bin and redshift correlation, see appendix\~A in \citealt{Koehlinger2016}.}.  The only approximation entering in the algorithm is the flat-sky approximation when introducing the Fourier decomposition of the spin-2 shear field.    This approximation is appropriate and accurate for current surveys as $\ell>80$  in all analyses published to date.
